### 2015-12-06 paper/paper.tex
Initial commit

%%%%%%%% BEGIN PREAMBLE %%%%%%%%

\documentclass[11pt]{article}
\usepackage{acl2015, latexsym, url, times, amssymb, amsmath, amsfonts}

%%%%%%%% END PREAMBLE %%%%%%%%
\title{Natural Language Processing\\
Fall 2015 Final Project:\\
Sense Disambiguation with the\\
Naive Bayes Classifier}
\author{Maximillian Dumas}

\begin{document}
\maketitle

\begin{abstract}
Beans beans beans. Right! This is a two column layout, so the title here is actually centered on the page.
\end{abstract}

\section{Introduction}
Introduce our solution to the problem and necessary background information.

\section{The Naive Bayes Classifier}
Explain in exact detail the Naïve Bayes Classifier algorithm. This should more or less follow the format of project proposal

\section{Implementation}
Discuss exactly what techniques were used to implement the algorithm. What technical issues/limitations resulted in changes? Figure out exactly what techniques we're co-opting for pos-tagging, lemmatizing, etc!

Talk about the baseline also.

Switched to POS-informed lemmatizer instead of simply choosing lemma with largest count, doubled accuracy.

\section{Analysis}
Discuss here the performance and results of applying the algorithm.

\section{Conclusion}
Discuss here what went wrong, what went right, avenues for improvement, further reading, etc.

\end{document}